Work Fragmentation in Developer Interaction Data

### 2016-01-29 paper_2.tex
RQ5 pass

Category \emph{perspective} deals with events related to perspective change in Eclipse, such as switching from the coding to the debugging perspective, or vice-versa \RR{other perspectives too}. This is more indicative of ... \RR{what?}

@@

Category \emph{refactoring} regroups usage of refactorings tools (rename, move methods, extract methods, etc). This is indicative of edition activities, rather than program comprehension. \RR{expand}

Category \emph{clean-build} regroup events related to build system usage (build the system, clean the build, run the build, the tests, etc). This is indicative of activities in which the programmer expects the program to be in a testable state. \RR{out of inspiration and time for now ...}

@@

After a positive interruption, we can see that the more common activities are text edition, refactoring, file management and, to a lesser extent, navigation around perspectives. On the other hand the navigation around classes, find and replace, searches, debug and clean and build are actions more common after a negative one. The navigation in the code and control version activities are more common after the recovery time, having a lower frequency during a positive o negative interruption.

We can conclude that after a positive interruption there are more assertive actions with a clear intention like refactoring a class or saving all changes; these activities are related to program edition. After a negative interruption, there are more uncertainty and doubt judging by the activities performed like debugging, searching for text or navigating around classes; these are activities related to program comprehension. These categories of events agree with our previous hypothesis that positive interruptions are related to information seeking \cite{PR11, LVD06}, while negative interruptions are related to actual task switches where context needs to be rebuilt afterwards.

It is noticeable the relation between type of events. For example, it is natural to think that if there are many debugging events there are also a lot of clean and build events, as they are necessary to begin a debugging session and in most cases it is automatically performed by the IDE. There is also a link between high level navigation and debugging because navigating into a method may take the programmer to another class.

>>>>

Category \emph{perspective} deals with events related to perspective change in Eclipse, such as switching from the coding to the debugging perspective, or vice-versa \RR{other perspectives too}. We have no expectation as to whether this kind of activity is more indicative of program understanding or program editing. \RR{or do we?}

@@

Category \emph{refactoring} regroups usage of refactorings tools (rename, move methods, extract methods, etc). This is indicative of edition activities, rather than program comprehension. \RR{expand?}

Category \emph{clean-build} regroup events related to build system usage (build the system, clean the build, run the build, the tests, etc). This is indicative of activities in which the programmer expects the program to be in an executable state. \RR{out of inspiration and time for now ...}

@@

After a positive interruptions the events belonging to the \emph{edit-text} category, the \emph{file} category, the \emph{refactoring} and the \emph{perspective} category, are more frequent than in the average. In addition, these categories of events are less frequent than the average after a negative interruptions. Of these four, the two categories with the largest differences between positive and negative interruptions are \emph{edit-text} and \emph{refactoring}. Both of these categories are much more clearly associated with edition activities than program comprehesion, and thus fit the previous hypothesis. Of the two other categories, \emph{file} management is also related to program edition activities, why the role of the \emph{perspective} category is still to clarify.

After a negative interruption, the \emph{debug}, \emph{search}, \emph{high-nav}, \emph{find-replace}, and \emph{clean-build} are all more common than average. Of note, the \emph{find-replace} category is also more common in positive interruptions. The three categories of \emph{debug}, \emph{search}, and \emph{high-nav} are clearly associated with program comprehension rather than program edition. The \emph{clean-build} category is not strongly associated with program comprehension, but it is however associated with debugging activities. The \emph{find-replace} category which is more common in both positive and negative interruptions can be, as explained above, both used for find-and-replace text editions, and also for conventional search, which could explain its presence in both kinds of interruptions.

Finally, the \emph{text-nav} and \emph{control} categories are less common in both positive and negative interruptions than in the average. We were somewhat surprised that \emph{text-nav} was less common in positive interruptions than the average, as we expected it to be associated with program edition behavior. We note that it is however more frequent in positive interruptions rather than negative interuptions. Additional exploration is necessary to understand the behavior of the \emph{control} category.

Summing up, we can conclude that edition-related activities are indeed more frequent after a positive interruption, while comprehension-related activities are more frequent after a negative interruption.  These categories of events agree with our previous hypothesis that positive interruptions are related to information seeking behavior: programmers would interrupt their IDE activities when encountering a roadblock, find the information elsewhere (from the internet, an expert, etc), and return to the IDE with an increased productivity. This also agrees with the literature \cite{PR11, LVD06}. The categories common after nagative interruptions  also agree with our hypothesis that negative interruptions are related to actual task switches where context needs to be rebuilt afterwards, an activity that involves program comprehension \RR{refs here}.

After a positive interruption, we can see that the more common activities are text edition, refactoring, file management and, to a lesser extent, navigation around perspectives. On the other hand the navigation around classes, find and replace, searches, debug and clean and build are actions more common after a negative one. The navigation in the code and control version activities are more common after the recovery time, having a lower frequency during a positive o negative interruption. Of course, this exploratory study does not allow us to confirm these hypotheses. However, the evidence we have discovered so far does not disprove them, which means that our confidence in these hypotheses is higher as a consequence.

%We can conclude that after a positive interruption there are more assertive actions with a clear intention like refactoring a class or saving all changes; these activities are related to program edition. After a negative interruption, there are more uncertainty and doubt judging by the activities performed like debugging, searching for text or navigating around classes; these are activities related to program comprehension.

%It is noticeable the relation between type of events. For example, it is natural to think that if there are many debugging events there are also a lot of clean and build events, as they are necessary to begin a debugging session and in most cases it is automatically performed by the IDE. There is also a link between high level navigation and debugging because navigating into a method may take the programmer to another class.
